The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2023-08-30 ms/ms.tex

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within land components of Earth System Models.
Understanding how these processes respond to CO$_2$ concentrations, and how they themselves impact CO$_2$ concentrations, is a critical facet of climate change research.  
Our certainty in climate model projections varies by domain and generally decreases with extended time horizons \cite{koven2022}.
Centennial-scale estimates of the cumulative terrestrial carbon sink have been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}. 
A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Effectively synthesizing the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns can help to parameterize increasingly comprehensive land modeling platforms.

Uncertainty in the model projections of the carbon cycle is manifest in both uncertainty around the appropriate algorithms and model structures used, and also in the parameterizations of these models. Model inter-comparison projects (MIPs) are the primary means by which structural uncertainty is assessed (refs). While these have had tremendous utility in capturing and assessing wide ranges of model assumptions, it can nonetheless be difficult to interpret the differences between models, or even between subsequent versions of the same model, due to the multiplicity of structural and parametric variations \cite{mcneall2016}. To allow for more tractable experimental design each model is typically allowed only a single parameterization of many plausible parameter sets. Thus, MIPs typically conflate parametric and structural uncertainty, and de-emphasize the consequences of uncertain model calibration in estimates of future land state trajectories.

@@

Several promising avenues for model calibration are in development (lavendar, musa, CES), many of which require the construction of large parameter perturbation ensembles (PPEs) \cite{qian2018}. 
PPEs have played an important role in assessing projection uncertainty in climate models \cite{murphy2004,sanderson2008,booth2012,hawkins2019,yamazaki2021,peatier2022,tett2022}.
PPEs also form a basis for automated model calibration through, for example, history matching \cite{williamson2013,williamson2017,hourdin2020,couvreux2021} or Bayesian uncertainty quantification \cite{cleary2021}.
A major goal of this project is to enable automated calibration of the Community Land Model (CLM) \cite{lawrence2019}. This extends the work of \citeA{dagon2020} which utilized a PPE to calibrate a subset of CLM parameters. Here we include a broader suite of parameters and continue to enhance the software tools for parameter perturbation and optimization.

@@

\section{Experiment Description}
\label{methods}
\subsection{Model description}
\label{sect:md}
This experiment utilizes the Community Land Model configuration version 5.1 (CLM5.1) of the Community Terrestrial Systems Model. The model source code and documentation are available online (\url{https://github.com/ESCOMP/CTSM}), as is a full model description \cite{lawrence2019}.

Relative to CLM5.0, version 5.1 includes minor bug fixes, parameter adjustments, and the implementation of biomass heat storage \cite{swenson2019}. The PPE experiment required additional code modifications to enable systematic variation of the full suite of model parameters (many of which were previously `hard-coded'). We utilized the biogeochemical cycling version of CLM, which includes active carbon and nitrogen cycles. This is as opposed to the satellite phenology mode, which is driven by observed canopy properties, and was explored in \citeA{dagon2020}. We used the model in land-only mode driven by observed meteorological forcing from the GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}), with the crop model turned off.

@@

To manage computational cost we leveraged the Matrix-CN spin-up mode recently implemented within CLM \cite{lu2020}. This new module utilizes a linearized simplification of CLM's biogeochemistry to significantly reduce spin-up time. Our spin-up protocol featured 20 years in accelerated decomposition mode (see \citeA{lawrence2019} for details), followed by 80 years of Matrix-CN, followed by 40 years of `normal' mode, cycling over a ten-year forcing dataset (described below). This protocol was designed to achieve sufficiently equilibrated model states, while minimizing computational time. This spin-up methodology did not always reach full equilibration of deep soil carbon (beyond 1 meter depth). Certain inferences about deep soil carbon from this ensemble are therefore subject to uncertainty due to spin-up concerns.

\subsection{Sparsegrid}
\label{sect:sg}
Another control on model cost is resolution. Most CLM simulations utilize nominal 1$^\circ$ resolution, which equates to about 20,000 land grid cells. In order to manage computational cost, parameter perturbation experiments often use lower resolution, such as 4$^\circ$x5$^\circ$ \cite{dagon2020}. Here, we use a clustering algorithm to achieve an alternative low resolution configuration.

@@

Given the large parameter space, to conduct an initial analysis of the response of the model to parametric uncertainty, we decided to vary each parameter independently, exploring the impact of low and high values. To define parameter ranges we created an online spreadsheet and solicited domain-area experts to provide a minimum and maximum value for each parameter. In some cases literature values were directly utilized, but in the many cases, expert judgment was used. The spreadsheet, with literature references and parameter descriptions is available online and in appendix zqz. In some cases parameters could not or should not be perturbed independently, because they feature inherent co-dependence. In the case of nitrogen fixation costs, we opted to perturb the parameters independently, but also as a group (`KCN'), to enforce a change in overall nitrogen limitation in lieu of switching from one uptake pathway to another. We likewise opted to perturb soil hydraulic parameters directly (e.g. saturated hydraulic conductivity) and also via perturbations to the soil texture (percent sand, clay, and silt).

Each simulation ran for 150 years, with the first 140 for spin-up (as described above in Section \ref{sect:mcn}), followed by a 10-year period for analysis. We opted for six different forcing scenarios to understand the intersection of parameter effects with different forcing agents associated with climate change (Table \ref{tab:exps}). The GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}) served as our atmospheric forcing, and is the default forcing data for CLM5 \cite{lawrence2019}. We applied climate and CO$_2$ anomalies independently, in order to disentangle their effects on parameter rankings. Future and pre-industrial climate forcing datasets were prepared by adding GSWP3v1 anomalies from 2005-2014 to the corresponding mean climate change signal (see Table \ref{tab:exps}). We inferred the mean climate change signal using the CESM2 large ensemble experiment \cite{rodgers2021}, computed as the average of the difference between the period of interest and present day for the six atmospheric forcing variables. Our future climate scenario utilizes the SSP3-7.0 forcing in order to allow comparison with the existing CESM2 large ensemble. Our software infrastructure has greatly reduced the effort and computational cost associated with ensemble generation, such that it would be easy to add additional ensembles, to explore other emission scenarios or climate regimes (e.g. Last Glacial Maximum).

\label{sect:exps}
 \begin{table}[h]
 \caption{Forcing Scenarios}
 \centering
 \begin{tabular}{l c c c c}
 \hline
  Name  & Meteorology & CO$_2$ (ppmv) & N addition & Description \\
 \hline
   CTL2010  & 2005-2014 & 367 & - & control experiment\\
   C285        & 2005-2014 & 285 & - & low CO$_2$ \\
   C867        & 2005-2014 & 867 & - & high CO$_2$ \\
   AF1855    & 1851-1860 & 367 & - & pre-industrial climate \\
   AF2095    & 2091-2100 & 367 & - & late century climate (SSP3-7.0) \\
   NDEP      & 2005-2014 & 367 & 5g/m$^2$ & enhanced nitrogen deposition \\
 \hline
 \end{tabular}
 \label{tab:exps}
 \end{table}

@@

Given the challenging size of the CLM5-BGC parameter space, and the need to probe and document parametric uncertainties within operational land surface models, a major goal of this project was to develop a fast configuration of CLM5-BGC that would allow for a large number of simulations given the available computational resources.
Combining the CN-Matrix spinup approach with our sparsegrid formulation (see Section \ref{sect:sg} for details) yielded a configuration approximately 500 times cheaper than the 1-degree configuration most often used for CLM simulations (Figure \ref{fig:sims}).
There are 22648, 5666, and 1764 land grid cells in standard 1$^{\circ}$, 2$^{\circ}$, and 4$^{\circ}$x5$^{\circ}$ CLM simulations, respectively, as compared to just 400 grid cells in the sparsegrid.
Likewise, whereas previous spin-up methodologies required 1500 years or longer to satisfy equilibrium criteria, the CN-matrix approach yielded satisfactory spin-up for our experiment within 140 years.

@@

Overall, we found substantial impacts of parametric uncertainty on model behaviour, in some cases exceeding the magnitude of climate scenario effects (Figure \ref{fig:ranges}). One perturbation, reducing the heat capacity of sand by 20\%, proved destructive in the future climate scenario, resulting in inhospitably hot soil conditions and widespread plant death. A perturbation of $\pm$20\% likely exceeds the reasonable range for this parameter, but this simulation was instructive for exposing the model's response to hot soils. Carrying out an extensive PPE increases the possibility of exposing unexpected model behavior, which may belie unforeseen tipping points, brittle parameterizations,  and/or bugs.

A small fraction of parameters tend to explain a large amount of the ensemble variance (Figure \ref{fig:variance}) for any given variable or metric. For example the same four parameters explain up to 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig \ref{supp:fcev}). Most of the parameter perturbations had a relatively small impact on any given output variable. As such, the effective parameter space dimensionality for any given CLM process is likely much smaller than would be expected from having 193 independent parameters, reflecting the fact that land surface models like CLM comprise representations of numerous interconnected domains. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/ranges.png}
\caption{Global annual gross primary production (GPP) and net ecosystem production (NEP) across six forcing scenarios. Circles mark the CLM5.1 default simulation, and the bars span the ensemble ranges.}
\label{fig:ranges}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{Cumulative fraction of variance explained by the most influential parameters on gross primary production and canopy evaporation. For example approximately 96\% of canopy evaporation variance is explained by the first ten parameters regardless of forcing, whereas between 74 and 92\% of GPP variance is explained by the first ten parameters. Only the top 50 parameters are shown (of 193) to improve the visualization. Note that parameter rankings differ by variable.}
\label{fig:variance}
\end{figure}

@@

There were several barriers to perturbing the full set of CLM5-BGC parameters. Firstly, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In this experiment we perturbed 193 parameters across a variety of CLM processes (Figure \ref{fig:params}). This is not the full set of CLM parameters, as some processes were not included, such as crops. In carrying out this experiment, we have likewise realized that the epistemology of climate model `parameters' is highly nuanced, and that defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM.

@@

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing allocation. As such, the computational cost of inferring parameter effects is a major constraint. We were able to reduce ensemble generation computational cost 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). As long as computational resources remain constrained, designing faster model configurations and more efficient sampling strategies will be important for effective model calibration and uncertainty quantification.

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each parameter. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That being said, the majority of parameter perturbations had small effects for any specific model variable or metric, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, for example, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be tractable parameter estimation sub-problems if global calibration of 193 parameters proves unattainable.

We opted for six forcing scenarios to identify parameter effects not just in present-day but in response to pre-industrial or future climate conditions. The parameters that were most influential under present-day conditions did not necessarily match the set of parameters controlling the response to future forcing (Figure \ref{fig:et}). We found that many acclimation parameters, which were not as important for determining present-day evapotranspiration (ET), were among the most influential on the response of ET to future climate. Similarly, the parameters controlling leaf area index varied significantly depending on biome (Figure \ref{fig:lai}).

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. A primary utility of our dataset is that we can diagnose parameter effects without the uncertainty contributed by an emulator or a regression model. As such, it is easy to diagnose which parameters are most influential on a given process. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now, in addition to seeing the definition of a given parameter, and the relevant equations, a model user can easily investigate the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}).

@@

\acknowledgments
Enter acknowledgments here. This section is to acknowledge funding, thank colleagues, enter any secondary affiliations, and so on.

>>>>

\section{Introduction}
Water availability, land temperature extremes, fire risk, and crop productivity will all see impacts from climate change, and are among the many processes represented within the terrestrial components of Earth System Models. Understanding how these processes respond to and influence CO$_2$ concentrations is a critical facet of climate change research. Our certainty in climate model projections varies by domain and generally decreases with extended time horizons \cite{koven2022}. Centennial-scale estimates of the cumulative terrestrial carbon sink have been especially challenging, with high uncertainty persisting across model generations \cite{friedlingstein2014,arora2020}. A portion of this uncertainty is irreducible, inherent to the challenge of predicting vegetation dynamics in a novel climate \cite{lovenduski2017}. Effectively synthesizing the ongoing expansion of observational data sources from remote sensing, meteorological stations, flux towers, and field campaigns can help to constrain increasingly comprehensive land modeling platforms.

Uncertainty in the model projections of the carbon cycle is manifest in both uncertainty around the appropriate algorithms and model structures used, and also in the parameterizations of these models (Luo and Schuur 2020). Model inter-comparison projects (MIPs) are the primary means by which structural uncertainty is assessed (refs). While these have had tremendous utility in capturing and assessing wide ranges of model assumptions, it can nonetheless be difficult to interpret the differences between models, or even between subsequent versions of the same model, due to the multiplicity of structural and parametric variations \cite{mcneall2016}. To allow for more tractable experimental design each model is typically allowed only a single parameterization (despite the existence of many plausible parameter sets) due to the cost of conducting multiple experiments with different parameterizations (e.g., in the case of the TRENDY land model intercomparison or CMIP). Thus, MIPs typically conflate parametric and structural uncertainty, and de-emphasize the consequences of uncertain model calibration in estimates of future land state trajectories.

@@

Several promising avenues for model calibration are in development (lavendar, musa, CES), many of which require the construction of large parameter perturbation ensembles (PPEs) \cite{qian2018}. 
PPEs have played an important role in assessing projection uncertainty in climate models \cite{murphy2004,sanderson2008,booth2012,hawkins2019,yamazaki2021,peatier2022,tett2022}.
PPEs also form a basis for automated model calibration through, for example, history matching \cite{williamson2013,williamson2017,hourdin2020,couvreux2021} or Bayesian uncertainty quantification \cite{cleary2021}. A parameter sensitivity analysis was conducted with ORCHIDEE using a semi-analytic spin-up (SASU) offered by a matrix approach to understand 34 biogeochemistry parameters according to variance explained by each parameter (Huang et al. 2018). 
A major goal of this project is to enable automated calibration of the Community Land Model (CLM) \cite{lawrence2019}. This extends the work of \citeA{dagon2020} which utilized a PPE to calibrate a subset of CLM parameters. Here we include a broader suite of parameters and continue to enhance the software tools for parameter perturbation and optimization.

@@

\section{Experiment Description}
\label{methods}
\subsection{Model description}
\label{sect:md}
This experiment utilizes the Community Land Model configuration version 5.1 (CLM5.1) of the Community Terrestrial Systems Model. The model source code and documentation are available online (\url{https://github.com/ESCOMP/CTSM}), as is a full model description \cite{lawrence2019}. (I again can't figure out how to comment, but this maybe involve a discussion on model naming (CTSM vs. CLM))

Relative to CLM5.0, version 5.1 includes minor bug fixes, parameter adjustments, and the implementation of biomass heat storage \cite{swenson2019}. (Birch et al. 2019) The PPE experiment required additional code modifications to enable systematic variation of the full suite of model parameters (many of which were previously `hard-coded', i.e., written directly as numerical values in the model source code). We utilized the biogeochemical cycling version of CLM, which includes active carbon and nitrogen cycles. This is as opposed to the satellite phenology mode, which is driven by observed canopy properties and thus has fewer internal feedbacks, and was explored in \citeA{dagon2020}. We used the model in land-only mode driven by observed meteorological forcing from the GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}), with the crop model turned off.

@@

To manage computational cost we leveraged the Semi-Analytic spin-up mode (i.e., SASU) recently implemented within CLM5 \cite{lu2020} (Liao et al. 2023). This new module utilizes a matrix representation on CLM's biogeochemistry to significantly reduce spin-up time by one order of magnitude (Liao et al. 2023, Luo et al. 2022). Our spin-up protocol featured 20 years in Accelerated Decomposition spin-up mode (i.e. AD) (see \citeA{lawrence2019} (Thornton  &  Rosenbloom, 2005) for details), followed by 80 years of SASU, followed by 40 years of Native Dynamic spin-up mode (i.e. ND), cycling over a ten-year forcing dataset (described below, Liao et al. 2023). This protocol was designed to achieve sufficiently equilibrated carbon and nitrogen states, while minimizing computational time. This spin-up methodology did not always reach full equilibration of deep soil carbon (beyond 1 meter depth) in very few points at high latitude. Certain inferences about deep soil carbon from this ensemble are therefore subject to uncertainty due to spin-up concerns.

\subsection{Sparse grid}
\label{sect:sg}
Another control on model cost is spatial resolution. Most global CLM simulations utilize nominal 1$^\circ$ resolution, which equates to about 20,000 land grid cells. In order to manage computational cost, parameter perturbation experiments often use lower resolution, such as 4$^\circ$x5$^\circ$ \cite{dagon2020}. Here, we use a clustering algorithm to achieve an alternative low resolution configuration.

@@

Given the large parameter space, to conduct an initial analysis of the response of the model to parametric uncertainty, we decided to vary each parameter independently, exploring the impact of low and high values. To define parameter ranges we created an online spreadsheet and solicited domain-area experts to provide a minimum and maximum value for each parameter. In some cases literature values were directly utilized, but in many cases, expert judgment was used. The spreadsheet, with literature references and parameter descriptions is available online and in appendix zqz. In some cases parameters could not or should not be perturbed independently, because they feature inherent co-dependence. In the case of nitrogen fixation costs, we opted to perturb the parameters independently, but also as a group (`KCN'), to enforce a change in overall nitrogen limitation in lieu of switching from one uptake pathway to another. We likewise opted to perturb soil hydraulic parameters directly (e.g. saturated hydraulic conductivity) and also via perturbations to the soil texture (percent sand, clay, and silt).

Each simulation ran for 150 years, with the first 140 for spin-up (as described above in Section \ref{sect:mcn}), followed by a 10-year period for analysis. We opted for six different forcing scenarios to understand the intersection of parameter effects with different forcing agents associated with climate change (Table \ref{tab:exps}). The GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}) served as our atmospheric forcing, and is the default forcing data for CLM5 \cite{lawrence2019}. We applied climate and CO$_2$ anomalies independently, in order to disentangle their effects on parameter rankings. Future and pre-industrial climate forcing datasets were prepared by adding GSWP3v1 anomalies from 2005-2014 to the corresponding mean climate change signal (see Table \ref{tab:exps}; AF2095 and AF1855). We inferred the mean climate change signal using the CESM2 large ensemble experiment \cite{rodgers2021}, computed as the average of the difference between the period of interest and present day for the six atmospheric forcing variables. Our future climate scenario utilizes the end-of-century climate from the SSP3-7.0, a high-emissions scenario, forcing in order to allow comparison with the existing CESM2 large ensemble. Our software infrastructure has greatly reduced the effort and computational cost associated with ensemble generation, such that it would be easy to add additional ensembles, to explore other emission scenarios or climate regimes (e.g. Last Glacial Maximum).

\label{sect:exps}
 \begin{table}[h]
 \caption{Forcing Scenarios}
 \centering
 \begin{tabular}{l c c c c}
 \hline
  Name  & Meteorology & CO$_2$ (ppmv) & N addition & Description \\
 \hline
   CTL2010  & 2005-2014 & 367 & - & control experiment\\
   C285        & 2005-2014 & 285 & - & low CO$_2$ \\
   C867        & 2005-2014 & 867 & - & high CO$_2$ \\
   AF1855    & 1851-1860 & 367 & - & pre-industrial climate \\
   AF2095    & 2091-2100 & 367 & - & late century climate (SSP3-7.0) \\
   NDEP      & 2005-2014 & 367 & 5g N m$^-2$  y$^-1$ & enhanced nitrogen deposition \\
 \hline
 \end{tabular}
 \label{tab:exps}
 \end{table}

@@

Given the challenging size of the CLM5-BGC parameter space, and the need to probe and document parametric uncertainties within operational land surface models, a major goal of this project was to develop a fast configuration of CLM5-BGC that would allow for a large number of simulations given the available computational resources.
Combining the SASU spinup approach with our sparsegrid formulation (see Section \ref{sect:sg} for details) yielded a configuration approximately 500 times cheaper than the 1-degree configuration most often used for CLM simulations (Figure \ref{fig:sims}).
There are 22648, 5666, and 1764 land grid cells in standard 1$^{\circ}$, 2$^{\circ}$, and 4$^{\circ}$x5$^{\circ}$ CLM simulations, respectively, as compared to just 400 grid cells in the sparsegrid.
Likewise, whereas previous spin-up methodologies required 1500 years or longer to satisfy equilibrium criteria, the SASU approach enabled by matrix representation yielded satisfactory spin-up efficiency for our experiment within 140 years.

@@

Overall, we found substantial impacts of parametric uncertainty on model behaviour, in some cases exceeding the magnitude of climate scenario effects (Figure \ref{fig:ranges}). One perturbation, reducing the heat capacity of sand by 20\%, proved destructive in the future climate scenario, resulting in inhospitably hot soil conditions and widespread plant death. A perturbation of $\pm$20\% likely exceeds the reasonable range for this parameter, but this simulation was instructive for exposing the model's response to hot soils. Carrying out an extensive PPE increases the possibility of exposing unexpected model behavior, including unforeseen tipping points, brittle parameterizations,  and/or bugs.

A small fraction of parameters tend to explain a large amount of the ensemble variance (Figure \ref{fig:variance}) for any given variable or metric. For example, the same four parameters explain up to 90\% of the canopy evaporation variance across all the forcing scenarios (Supp Fig \ref{supp:fcev}). In general,most of the parameter perturbations had a relatively small impact on any given output variable. As such, the effective parameter space dimensionality for any given CLM process is likely much smaller than would be expected from having 193 independent parameters, reflecting the fact that land models like CLM comprise representations of numerous interconnected domains. While the principle of land models is that these are all interlinked via complex feedback processes, many parameters nonetheless have impacts that are mostly limited to their own domain.

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/ranges.png}
\caption{Global annual gross primary production (GPP) and net ecosystem production (NEP) across six forcing scenarios. Circles mark the CLM5.1 default simulation, and the bars span the ensemble ranges generated by individual parameter perturbations.}
\label{fig:ranges}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{../figs/variance.pdf}
\caption{Cumulative fraction of variance explained by the most influential parameters on gross primary production and canopy evaporation. For example, approximately 96\% of canopy evaporation variance is explained by the first ten parameters regardless of forcing, whereas between 74 and 92\% of GPP variance is explained by the first ten parameters. Only the top 50 parameters are shown (of 193) to improve the visualization. Note that parameter rankings differ by variable.}
\label{fig:variance}
\end{figure}

@@

There were several barriers to perturbing the full set of CLM5-BGC parameters. First, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In this experiment we perturbed 193 parameters across a variety of CLM processes (Figure \ref{fig:params}). This is not the full set of CLM parameters, as some processes were not included, such as crop phenology and management. In carrying out this experiment, we have likewise realized that the epistemology of climate model `parameters' is highly nuanced, and that defining the parameters within a given model structure can be somewhat subjective, such that it would be difficult to collate a comprehensive or definitive set of parameters for a model like CLM.

@@

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response is potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing allocation. As such, the computational cost of inferring parameter effects is a major constraint. We were able to reduce ensemble generation computational cost 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a Semi-Analytic Spin-Up (SASU) offered by the matrix approach to land biogeochemistry models (Lu et al. 2020, Luo 2022, Liao et al. 2023) and a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). As long as computational resources remain constrained, designing faster model configurations and more efficient sampling strategies will be important for effective model calibration and uncertainty quantification.

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each parameter. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That said, the majority of parameter perturbations had small effects for any specific model variable or metric, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, for example, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be tractable parameter estimation sub-problems if global calibration of 193 parameters proves unattainable.

We opted for six forcing scenarios to identify parameter effects not just in present-day but in response to pre-industrial or future climate conditions. The parameters that were most influential under present-day conditions did not necessarily match the set of parameters controlling the response to future forcing (Figure \ref{fig:et}). We found that many acclimation parameters, which were not as important for determining present-day evapotranspiration (ET), were among the most influential on the response of ET to future climate. Similarly, the parameters controlling leaf area index varied significantly depending on biome (Figure \ref{fig:lai}). This emphasizes the importance of testing models not only under mean-state conditions, but also under experimentally perturbed conditions under global change analogues \cite(Wieder:2019).

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. A primary utility of our dataset is that we can diagnose parameter effects without the uncertainty contributed by an emulator or a regression model. As such, it is easy to diagnose which parameters are most influential on a given process. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now, in addition to seeing the definition of a given parameter, and the relevant equations, a model user can easily investigate the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}). This could be useful for investigating obvious model deficiencies, such as regional LAI biases, or to understand potential model responses under different forcing scenarios (e.g. plant survival under low CO2 conditions).

@@

\acknowledgments
Enter acknowledgments here. This section is to acknowledge funding, thank colleagues, enter any secondary affiliations, and so on. This material is based upon work supported by the National Center for Atmospheric Research (NCAR), which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977. Computing and data storage resources, including the Cheyenne supercomputer (doi:10.5065/D6RX99HX) were provided by the Climate Simulation Laboratory at NCAR’s Computational and Information Systems Laboratory (CISL). FL is supported by the National Key Research and Development Program of China (2022YFE0106500). CDK acknowledges support by the Director, Office of Science, Office of Biological and Environmental Research of the US Department of Energy under contract DE-AC02-05CH11231 through the Regional and Global Model Analysis Program (RUBISCO SFA).
